---
title: "2026 EBS pollock assessment dashboard"
subtitle: "Schedule, data readiness, bridge progress, and review status"
format:
  dashboard:
    theme: cosmo
    orientation: rows
    scrolling: true
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

candidate_dirs <- c(".", file.path("assessment-years", "2026"))
year_dir <- candidate_dirs[
  file.exists(file.path(candidate_dirs, "data-manifest.csv"))
][1]

if (is.na(year_dir)) {
  stop("Could not locate the 2026 assessment manifests.")
}

read_manifest <- function(filename) {
  read.csv(
    file.path(year_dir, filename),
    check.names = FALSE,
    stringsAsFactors = FALSE,
    na.strings = c("", "NA")
  )
}

data_manifest <- read_manifest("data-manifest.csv")
run_manifest <- read_manifest("run-manifest.csv")
review_manifest <- read_manifest("review-decisions.csv")
deliverables <- read_manifest("deliverables.csv")
schedule <- read_manifest("schedule.csv")

candidate_data <- sum(data_manifest$availability_status == "candidate_available")
pending_data <- sum(data_manifest$availability_status == "pending")
validated_data <- sum(data_manifest$availability_status == "validated")
open_decisions <- sum(!review_manifest$status %in% c("complete", "final", "ssc_accepted"))

status_order <- c(
  "validated", "candidate_available", "active", "working", "planned",
  "pending", "incomplete", "missing", "not_started"
)
status_colors <- c(
  validated = "#2b8c4b",
  candidate_available = "#d99a00",
  active = "#2171b5",
  working = "#2171b5",
  planned = "#6b7280",
  pending = "#d97706",
  incomplete = "#b45309",
  missing = "#b91c1c",
  not_started = "#6b7280"
)

format_table <- function(x, columns, labels = columns) {
  out <- x[, columns, drop = FALSE]
  names(out) <- labels
  knitr::kable(out, format = "html", escape = TRUE, row.names = FALSE)
}
```

## How to use this dashboard {height=230}

### What this page shows

::: {.callout-note}
This is a **read-only summary** of five CSV tracking files in this assessment-year
folder. Edit those files—not the boxes or tables on this page—and the dashboard
updates the next time Quarto renders the project.

The inventory is current as of **July 17, 2026**. **Located; needs checking**
means a possible input file has been found, but its owner, version, units,
processing, and suitability still need confirmation. **Ready for assessment use**
means those checks are complete. **Awaiting delivery** means the expected input
has not yet been received. **Open decisions** are unresolved choices recorded in
the review tracker.
:::

## Assessment status {height=180}

```{r}
#| content: valuebox
#| title: "Located; needs checking"

list(icon = "database-check", color = "warning", value = candidate_data)
```

```{r}
#| content: valuebox
#| title: "Ready for assessment use"

list(icon = "check-circle", color = "success", value = validated_data)
```

```{r}
#| content: valuebox
#| title: "Awaiting delivery"

list(icon = "hourglass-split", color = "secondary", value = pending_data)
```

```{r}
#| content: valuebox
#| title: "Open decisions"

list(icon = "question-circle", color = "danger", value = open_decisions)
```

## Schedule {height=350}

### Review and delivery milestones

```{r}
format_table(
  schedule,
  c("milestone", "target_date", "purpose", "status"),
  c("Milestone", "Target", "Purpose", "Status")
)
```

## Data readiness {height=500}

### Status summary

```{r}
#| fig-alt: "Bar chart counting 2026 assessment data products by availability status."

data_counts <- table(data_manifest$availability_status)
data_counts <- data_counts[order(match(names(data_counts), status_order))]
barplot(
  data_counts,
  horiz = TRUE,
  las = 1,
  col = unname(status_colors[names(data_counts)]),
  border = NA,
  xlab = "Number of data products",
  main = "Data readiness"
)
```

### Items needing action

```{r}
data_actions <- data_manifest[
  data_manifest$availability_status != "validated",
  c("data_stream", "terminal_year", "availability_status", "validation_needed", "target_milestone")
]
format_table(
  data_actions,
  names(data_actions),
  c("Data stream", "Terminal year", "Status", "Next check", "Target")
)
```

## Bridge and model runs {height=420}

### Planned run sequence

```{r}
format_table(
  run_manifest,
  c("run_id", "role", "parent_run", "review_use", "author_status"),
  c("Run", "Role", "Parent", "Purpose", "Status")
)
```

### Important gap

::: {.callout-warning title="Unresolved bridge source"}
The exact September 2025 post-CIE draft and model configuration still need to be
identified. The 2024 September document is not an acceptable substitute.
:::

## Decisions and review {height=350}

### Choices that still need an owner or resolution

```{r}
review_actions <- review_manifest[
  !review_manifest$status %in% c("complete", "final", "ssc_accepted"),
  c("stage", "decision", "owner", "status")
]
format_table(
  review_actions,
  names(review_actions),
  c("Stage", "Decision needed", "Owner", "Status")
)
```

## Deliverables {height=450}

### Status summary

```{r}
#| fig-alt: "Bar chart counting 2026 assessment deliverables by status."

deliverable_counts <- table(deliverables$status)
deliverable_counts <- deliverable_counts[
  order(match(names(deliverable_counts), status_order))
]
barplot(
  deliverable_counts,
  horiz = TRUE,
  las = 1,
  col = unname(status_colors[names(deliverable_counts)]),
  border = NA,
  xlab = "Number of deliverables",
  main = "Deliverable status"
)
```

### Required outputs and checks

```{r}
format_table(
  deliverables,
  c("description", "validation", "status"),
  c("Deliverable", "Required check", "Status")
)
```

## Updating the dashboard {height=360}

### Source files and rebuild commands

Edit the source records in `assessment-years/2026/`:

- `schedule.csv` for dates and milestones;
- `data-manifest.csv` for deliveries and validation;
- `run-manifest.csv` for bridge and model progress;
- `review-decisions.csv` for decisions and dispositions; and
- `deliverables.csv` for document and output readiness.

Commit and push the changes. GitHub Actions validates the records and renders the
site automatically. To check the same changes locally before pushing, run:

```sh
Rscript scripts/validate_manifests.R
quarto render
```
